Title Goes Here

### 2021-04-06 analysis/paper/shceof.Rmd
Justifica latitudes de ozono

o3wave_lats <- c(-70, -45)
```

@@

(ref:t-vertical-cap) Regression coefficients between cEOF1 and zonal anomalies of mean air temperature (shaded) and ozone mixing ratio (contours, negative contours with dashed lines, labels in parts per billion by mass) averaged between 80ºS and 35ºS for the period 1979 -- 2019.

@@

```{r t-vertical, fig.cap = "(ref:t-vertical-cap)", fig.height=2.5, fig.width=width_column}
t %>% 
  .[lat %between% c(-80, -35)] %>% 
  .[, .(hgt = weighted.mean(hgt, cos(lat*pi/180)),
        o3 = weighted.mean(o3, cos(lat*pi/180)),
        t = weighted.mean(t, cos(lat*pi/180))), by = .(lev, time, lon)] %>%  
  melt(id.vars = c("lev", "time", "lon")) %>% 
  .[, value := Anomaly(value), by = .(time, lev, variable)] %>%
  .[cut(ceof$eof[[1]], 1)$left, on = "time"] %>% 
  sep_ReIm(format = "wider") %>% 
  .[, ":="(Real = as.numeric(scale(Real)), Imaginary = as.numeric(scale(Imaginary))), by = .(cEOF)] %>% 
  .[, FitLm(value, Real, Imaginary, se = TRUE), by = .(lev, lon, variable)] %>% 
  rm_intercept() %>% 
  .[, term := factor_ReIm(term)] %>% 
  # .[, z := metR:::standard_atmosphere$altitude(lev*100)] %>% 
  # .[, dz1 := c(NA, diff(estimate)/diff(z)) , by = .(term, lon, variable)] %>% 
  # .[, dz2 := c(diff(estimate)/diff(z), NA) , by = .(term, lon, variable)] %>% 
  # .[, dz := (dz1 + dz2)/2] %>% 
  # .[, p.value := Pvaluate(estimate, std.error, df, "fdr"), by = .(term, variable)] %>%
  # .[lev < 900] %>% 
  ggplot(aes(lon, lev)) +
  geom_contour_fill(data = ~.x[variable == "t"],
                    aes(z = estimate, fill = ..level..),
                    breaks = AnchorBreaks(0, binwidth = 0.25, exclude = 0)) +
  geom_contour_tanaka(data = ~.x[variable == "t"],
                      aes(z = estimate),
                      breaks = AnchorBreaks(0, binwidth = 0.25, exclude = 0)) +
  geom_contour2(data = ~.x[variable == "o3"], 
                aes(z = estimate, linetype = factor(-sign(..level..))), 
                breaks = AnchorBreaks(0, exclude = 0)) +
  geom_text_contour(data = ~.x[variable == "o3"], 
                aes(z = estimate*1e9, label = signif(..level.., 2)),
                size = 2,
                skip = 1, stroke = 0.15, stroke.colour = "white", 
                label.placement = lable_placement_min(),
                breaks = AnchorBreaks(0, exclude = 0)) +

@@

The vertical distribution of these patterns is shown in Figure\ \@ref(fig:t-vertical)) that shows the regression coefficientes between cEOF1 and zonal anomalies of air temperature and zonal anomalies of ozone mixing ratio averaged between 80ºS and 35ºS. Temperature zonal anomalies associated with cEOF1 show a clear wave 1 pattern for both real and imaginary components throughout the atmosphere above 250 hPa  with a change in sign above 10 hPa. Following hydrostatic balance, this is the level in which the geopotential anomaly have maximum amplitude (not shown).

@@

```{r o3-regr, fig.cap = "(ref:o3-regr-cap)", fig.height=3, fig.width = width_column}
o3_regr %>% 
  periodic(lon = c(0, 360)) %>% 
  .[, p.value := Pvaluate(estimate, std.error, df, "fdr"), by = .(cEOF, term)] %>% 
  ggplot(aes(lon, lat)) +
  geom_contour_fill(aes(z = estimate, fill = ..level..),
                    breaks = AnchorBreaks(0, exclude = 0)) +
  geom_contour2(data = periodic(o3_mean, lon = c(0, 360)),
                aes(z = toc_z, linetype = factor(-sign(..level..))), 
                breaks = AnchorBreaks(0, exclude = 0), 
                bins = 5) +
  geom_contour_tanaka(aes(z = estimate), breaks = AnchorBreaks(0, exclude = 0)) +
  geom_contour_pval(aes(z = p.value)) +

  geom_qmap(~.x[lat <= -20]) +
  scale_fill_divergent_discretised("Total Ozone Column", 
                                   guide = guide_colorsteps_bottom(15)) + 
  scale_x_longitude() +
  scale_y_latitude() +
  scale_linetype(guide = "none") +
  coord_polar() +
  facet_grid(cEOF~term) +
  axis_labs_smol +
  tag_facets("panel")

@@

Indeed, Figure\ \@ref(fig:wave1-o3) shows the relationship between amplitude and phase of the planetary wave 1 in Total Ozone Column between 70ºS and 45ºS \textcolor{red}{(Habría que justifica mejor estas elecciones de bandas latitudinales a paritr del gráfico anterior)} and amplitude and phase of the EOF1. The correlation between the zonal wave 1 of ozone and EOF1 (computed as the mean cosine of the difference in phase) is `r wave_correlation$text`. As expected from the location of anomalies in Figure\ \@ref(fig:wave1-o3), the Real EOF1 drives the relationship with amplitude and the Imaginary EOF1 drives the relationship with phase (not shown).

>>>>

o3wave_lats <- c(-75, -45)
```

@@

(ref:t-vertical-cap) Regression coefficients between cEOF1 and zonal anomalies of mean air temperature (shaded) and ozone mixing ratio (contours, negative contours with dashed lines, labels in parts per billion by mass) averaged between `r knitr::combine_words(LatLabel(o3wave_lats))` for the period 1979 -- 2019.

@@

```{r t-vertical, fig.cap = "(ref:t-vertical-cap)", fig.height=2.5, fig.width=width_column}
t %>% 
  .[lat %between% o3wave_lats] %>% 
  .[, .(hgt = weighted.mean(hgt, cos(lat*pi/180)),
        o3 = weighted.mean(o3, cos(lat*pi/180)),
        t = weighted.mean(t, cos(lat*pi/180))), by = .(lev, time, lon)] %>%  
  melt(id.vars = c("lev", "time", "lon")) %>% 
  .[, value := Anomaly(value), by = .(time, lev, variable)] %>%
  .[cut(ceof$eof[[1]], 1)$left, on = "time"] %>% 
  sep_ReIm(format = "wider") %>% 
  .[, ":="(Real = as.numeric(scale(Real)), Imaginary = as.numeric(scale(Imaginary))), by = .(cEOF)] %>% 
  .[, FitLm(value, Real, Imaginary, se = TRUE), by = .(lev, lon, variable)] %>% 
  rm_intercept() %>% 
  .[, term := factor_ReIm(term)] %>% 
  # .[, z := metR:::standard_atmosphere$altitude(lev*100)] %>% 
  # .[, dz1 := c(NA, diff(estimate)/diff(z)) , by = .(term, lon, variable)] %>% 
  # .[, dz2 := c(diff(estimate)/diff(z), NA) , by = .(term, lon, variable)] %>% 
  # .[, dz := (dz1 + dz2)/2] %>% 
  # .[, p.value := Pvaluate(estimate, std.error, df, "fdr"), by = .(term, variable)] %>%
  # .[lev < 900] %>% 
  ggplot(aes(lon, lev)) +
  geom_contour_fill(data = ~.x[variable == "t"],
                    aes(z = estimate, fill = ..level..),
                    breaks = AnchorBreaks(0, binwidth = 0.25, exclude = 0)) +
  geom_contour_tanaka(data = ~.x[variable == "t"],
                      aes(z = estimate),
                      breaks = AnchorBreaks(0, binwidth = 0.25, exclude = 0)) +
  geom_contour2(data = ~.x[variable == "o3"], 
                aes(z = estimate, linetype = factor(-sign(..level..))), 
                breaks = AnchorBreaks(0, exclude = 0)) +
  geom_text_contour(data = ~.x[variable == "o3"], 
                aes(z = estimate*1e9, label = signif(..level.., 2)),
                size = 2,
                skip = 1, stroke = 0.15, stroke.colour = "white", 
                label.placement = lable_placement_min(),
                breaks = AnchorBreaks(0, exclude = 0)) +

@@

The vertical distribution of these patterns is shown in Figure\ \@ref(fig:t-vertical)) that shows the regression coefficientes between cEOF1 and zonal anomalies of air temperature and zonal anomalies of ozone mixing ratio averaged between `r knitr::combine_words(LatLabel(o3wave_lats))`. Temperature zonal anomalies associated with cEOF1 show a clear wave 1 pattern for both real and imaginary components throughout the atmosphere above 250 hPa  with a change in sign above 10 hPa. Following hydrostatic balance, this is the level in which the geopotential anomaly have maximum amplitude (not shown).

@@

```{r o3-regr, fig.cap = "(ref:o3-regr-cap)", fig.height=3, fig.width = width_column}
o3_regr %>% 
  periodic(lon = c(0, 360)) %>% 
  .[, p.value := Pvaluate(estimate, std.error, df, "fdr"), by = .(cEOF, term)] %>% 
  ggplot(aes(lon, lat)) +
  geom_contour_fill(aes(z = estimate, fill = ..level..),
                    breaks = AnchorBreaks(0, exclude = 0)) +
  geom_contour_tanaka(aes(z = estimate), breaks = AnchorBreaks(0, exclude = 0)) +
  geom_contour2(data = periodic(o3_mean, lon = c(0, 360)),
                aes(z = toc_z, linetype = factor(-sign(..level..))), 
                breaks = AnchorBreaks(0, exclude = 0), 
                bins = 5) +

  geom_contour_pval(aes(z = p.value)) +
  geom_hline(yintercept = o3wave_lats, size = 0.3, alpha = 0.5) +
  geom_qmap(~.x[lat <= -20]) +
  scale_fill_divergent_discretised("Total Ozone Column", 
                                   guide = guide_colorsteps_bottom(15)) + 
  scale_x_longitude() +
  scale_y_latitude() +
  scale_linetype(guide = "none") +
  coord_polar() +

  facet_grid(cEOF~term) +
  axis_labs_smol +
  tag_facets("panel")

@@

Indeed, Figure\ \@ref(fig:wave1-o3) shows the relationship between amplitude and phase of the cEOF1 and the amplitude and phase of the planetary wave 1 in Total Ozone Column between `r knitr::combine_words(LatLabel(o3wave_lats))`; as shown in Figure\ \@ref(fig:o3-regr), this is a wide latitude band that captures most of the area with significant regression coefficients. The correlation between the zonal wave 1 of ozone and EOF1 (computed as the mean cosine of the difference in phase) is `r wave_correlation$text`. As expected from the location of anomalies in Figure\ \@ref(fig:wave1-o3), the Real EOF1 drives the relationship with amplitude and the Imaginary EOF1 drives the relationship with phase (not shown).
